Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph for Clustering and Parameter Estimation

### 2021-02-19 summary.Rmd
fixed some of the proofs

\begin{equation}
  P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top
\end{equation}

@@

\begin{equation}
  P = X I_{pq} X^\top
\end{equation}

@@

\begin{equation}
\min_{c_i} \frac{1}{2} ||x_i - X_{-i} c_i||^2_2 + \lambda ||c_i||_1
\end{equation}

@@

\begin{equation}
  \max\limits_{i \in \{1, ..., n\}} ||Q_n \hat{z}_i^{(n)} - z_i^{(n)}|| = 
  O_P\bigg(\frac{(\log n)^c}{n^{1/2}} \bigg)
\end{equation}

@@

\begin{equation}
  \max\limits_{i, j} ||(\hat{v}^{(n)}_i)^\top \hat{v}_j^{(n)}|| = 
    O_P\bigg( \frac{(\log n)^c}{n \sqrt{\rho_n}}\bigg)
\end{equation}

@@

\begin{equation}
\max_{k, l \in \{1, ..., K\}} 
||\hat{\lambda}^{(kl)} - \lambda^{(kl)}|| = 
O_P \bigg(\frac{(\log n_k)^c}{\sqrt{n_k}} \bigg)
\end{equation}

@@

\begin{equation}
X = \blockdiag(\Lambda^{(1)}, ..., \Lambda^{(K)}) \in \mathbb{R}^{n \times K^2}
\end{equation}

@@

\begin{equation}
\Pi = U I_{K (K + 1) / 2, K (K - 1) / 2} U^\top
\end{equation}

@@

\begin{equation}
P = X U I_{p, q} (X U)^\top
\end{equation}

\begin{equation}
p = K (K + 1) / 2
\end{equation}

\begin{equation}
q = K (K - 1) / 2
\end{equation}

@@

**Lemma 3.1.** 
Let $P = V D V^\top$ be the spectral decomposition of the edge probability 
matrix for a PABM. Then $V V^\top = X (X^\top X)^{-1} X^\top$ where $X$ is 
defined as in equation (7).

*Proof of Lemma 3.1*. 
By Theorem 2, $P = X U I_{p, q} U^\top X^\top$, where $X$ is defined as in 
equation (7) and $p$ and $q$ are defined as in equations (10) and (11).
Alternatively, the spectral decomposition can be written as 
$P = V D V^\top = V |D|^{1/2} I_{p, q} |D|^{1/2} V^\top$ for the same $(p, q)$ 
and $|\cdot|^{1/2}$ is applied entry-wise. Thus for some 
$Q \in \mathbb{O}(p, q)$,

@@

\begin{equation}
\hat{r}_n^{(k)} \stackrel{p}{\to} r_n^{(k)}
\end{equation}

*Proof of Lemma 5.1*. 
This falls directly out of Theorem 5 of \citet{rubindelanchy2017statistical}. 
Since $\hat{V}_n^{(k)} \stackrel{p}{\to} V_n^{(k)}$ and $r(\cdot)$ is smooth, 
$\hat{r}_n^{(k)} \stackrel{p}{\to} r_n^{(k)}$.

**Lemma 5.2**. 
Let $\mu_n^{(k)} = \mu(V_n^{(k)})$ be the subspace incoherence of $V_n^{(k)}$ 
as defined by \citet{soltanolkotabi2012} (also note that 
$\mu(\hat{V}_n^{(k)}) = \mu(V_n^{(k)})$ using the extension of this definition 
by \citet{pmlr-v28-wang13}).

\begin{equation}
\mu_n^{(k)} = 0
\end{equation}

with probability 1 $\forall n, k$.

@@

**Lemma 5.3**. 
Let $(v_n^{(i)})^\top$ and $(\hat{v}_n^{(i)})^\top$ be the rows of $V_n$ and
$\hat{V}_n$ respectively. By \cite{rubindelanchy2017statistical},

\begin{equation}
\delta_n = \max_i ||\hat{v}_n^{(i)} - v_n^{(i)}|| \stackrel{p}{\to} 0
\end{equation}

@@

By (12) and (13),

\begin{equation}
r(\hat{V}_n^{(k)}) - \mu(\hat{V}_n^{(k)}) = r_n^{(k)} \stackrel{p}{\to} r_n^{(k)} > 0
\end{equation}

Then by Theorem 6 of \citet{pmlr-v28-wang13}, (14) and (15) imply that  
$\Pr(\lim\limits_{n \to \infty} \text{ SSC detection property holds for } \hat{V}_n) = 1 - \epsilon$.

>>>>

\begin{equation} \label{eq:pabm}
  P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top
\end{equation}

@@

\begin{equation} \label{eq:grdpg}
  P = X I_{pq} X^\top
\end{equation}

@@

\begin{equation} \label{eq:ssc}
\min_{c_i} \frac{1}{2} ||x_i - X_{-i} c_i||^2_2 + \lambda ||c_i||_1
\end{equation}

@@

\begin{equation} \label{eq:thm3}
  \max\limits_{i \in \{1, ..., n\}} ||Q_n \hat{z}_i^{(n)} - z_i^{(n)}|| = 
  O_P\bigg(\frac{(\log n)^c}{n^{1/2}} \bigg)
\end{equation}

@@

\begin{equation} \label{eq:thm4}
  \max\limits_{i, j} ||(\hat{v}^{(n)}_i)^\top \hat{v}_j^{(n)}|| = 
    O_P\bigg( \frac{(\log n)^c}{n \sqrt{\rho_n}}\bigg)
\end{equation}

@@

\begin{equation} \label{eq:thm6}
\max_{k, l \in \{1, ..., K\}} 
||\hat{\lambda}^{(kl)} - \lambda^{(kl)}|| = 
O_P \bigg(\frac{(\log n_k)^c}{\sqrt{n_k}} \bigg)
\end{equation}

@@

\begin{equation} \label{eq:xy}
X = \blockdiag(\Lambda^{(1)}, ..., \Lambda^{(K)}) \in \mathbb{R}^{n \times K^2}
\end{equation}

@@

\begin{equation} \label{eq:permutation}
\Pi = U I_{K (K + 1) / 2, K (K - 1) / 2} U^\top
\end{equation}

@@

\begin{equation} \label{eq:pabm-grdpg}
P = X U I_{p, q} (X U)^\top
\end{equation}

\begin{equation} \label{eq:p}
p = K (K + 1) / 2
\end{equation}

\begin{equation} \label{eq:q}
q = K (K - 1) / 2
\end{equation}

@@

**Lemma 3.1.** 
Let $P = V D V^\top$ be the spectral decomposition of the edge probability 
matrix for a PABM. Then $V V^\top = X (X^\top X)^{-1} X^\top$ where $X$ is 
defined as in (\ref{eq:xy}).

*Proof of Lemma 3.1*. 
By Theorem 2, $P = X U I_{p, q} U^\top X^\top$, where $X$ is defined as in 
(\ref{eq:xy}) and $p$ and $q$ are defined as in equations (\ref{eq:p}) and
(\ref{eq:q}). Alternatively, the spectral decomposition can be written as 
$P = V D V^\top = V |D|^{1/2} I_{p, q} |D|^{1/2} V^\top$ for the same $(p, q)$ 
and $|\cdot|^{1/2}$ is applied entry-wise. Thus for some 
$Q \in \mathbb{O}(p, q)$,

@@

\begin{equation} \label{eq:r-conv}
\hat{r}_n^{(k)} \stackrel{p}{\to} r_n^{(k)}
\end{equation}

*Proof of Lemma 5.1*. 
This falls directly out of Theorem 5 of \citet{rubindelanchy2017statistical}. 
Since $\hat{V}_n^{(k)} \stackrel{p}{\to} V_n^{(k)}$ and $r(\cdot)$ is 
continuous, $r(\hat{V}_n^{(k)}) \stackrel{p}{\to} r(V_n^{(k)})$.

**Lemma 5.2**. 
Let $\mu_n^{(k)} = \mu(V_n^{(k)})$ be the subspace incoherence of $V_n^{(k)}$ 
as defined by \citet{soltanolkotabi2012}. Then 
$\mu_n^{(k)} = 0$ 
with probability 1 $\forall n, k$.

@@

**Lemma 5.3**. 
Let $\hat{\mu}_n^{(k)} = \mu(\hat{V}_n^{(k)})$ using the extended definition of
$\mu(\cdot)$ from \citet{pmlr-v28-wang13}. Then

\begin{equation} \label{eq:mu-conv}
\hat{\mu}_n^{(k)} \stackrel{p}{\to} 0
\end{equation}

*Proof of Lemma 5.3*. 
This falls directly out of Theorem 5 of \citet{rubindelanchy2017statistical}. 
Since $\hat{V}_n^{(k)} \stackrel{p}{\to} V_n^{(k)}$ and $\mu(\cdot)$ is 
continuous, $\mu(\hat{V}_n^{(k)}) \stackrel{p}{\to} \mu(V_n^{(k)}) = 0$.

**Lemma 5.4**. 
Let $(v_n^{(i)})^\top$ and $(\hat{v}_n^{(i)})^\top$ be the rows of $V_n$ and
$\hat{V}_n$ respectively. By \citet{rubindelanchy2017statistical},

\begin{equation} \label{eq:delta-conv}
\delta_n = \max_i ||\hat{v}_n^{(i)} - v_n^{(i)}|| \stackrel{p}{\to} 0
\end{equation}

@@

By (\ref{eq:r-conv}) and (\ref{eq:mu-conv}),

\begin{equation} \label{eq:thm5-proof}
r(\hat{V}_n^{(k)}) - \mu(\hat{V}_n^{(k)}) \stackrel{p}{\to} 
r_n^{(k)} - \mu_n^{(k)} = r_n^{(k)} > 0
\end{equation}

for each $k = 1, ..., K$.

Then by Theorem 6 of \citet{pmlr-v28-wang13}, (\ref{eq:delta-conv}) and
(\ref{eq:thm5-proof}) imply that $\forall \epsilon > 0$, $\exists N > 0$ such 
that $\forall n > N$, 
$\Pr(\text{SSC detection property holds for } \hat{V}_n) = 1 - \epsilon$.
